Multi-Modal Modular Sensorimotor Contingencies for Deformable Objects

### 2016-08-29 twostagedo.tex
Finished learning.

The contour of the deformable object is used to obtain the ground truth for the training phase, which is the area enclosed by the contour. Geometrically, the contour is represented by a polygon with hundreds of sides, which can also be called a \textit{linear spline}.  The tracking of the deformable target in the ground truth has two stages:
\begin{enumerate}
 \item The target object is segmented from the environment.  The Canny algorithm is used to extract the edges of the image.  \comment{\sout{Next, a simple colour segmentation is used to remove edges that are not likely to belong to the object of interest.}  Since the object of interest has an outstanding colour, the hue of the pixels on every edge }
 \item The internal representation of its contour, the linear snake, is adapted to its new shape.  The algorithm assumes that no occlusions take place and that the contour does not bend over itself.
\end{enumerate}
For the first frame, the snake is initialised as a rectangle around the target. This rectangle fragments itself into smaller pieces. Each new vertex adheres itself to the closest point on an edge, as reached by growing rings of pixels around its current position, up to a certain threshold distance [\fref{fig:scene}(a)].  If there is no edge within that area, the control point remains were it is. This criterion was selected to make the representation more robust against edges that are easier to detect on some frames than in others.  Once the edge reappears, the control point adapts itself again.  Also, the distance between two consecutive control points is kept within the interval $[min\_distance, max\_distance]$ and two adjacent edges do not form angles smaller than $min\_angle$.  See \alref{alg:regularise}.  This allows the snake to adapt very quickly to the complexity of the contour it represents.

@@

This measurement is used in the computer vision community to evaluate segmentation algorithms.  $F=1$ means that there were no false predictions ($FN+FP = 0$).  $F \rightarrow 0$\footnote{This is read: the value of $F$ goes to zero or $F$ tends to zero.}, if there are no good predictions ($TP\rightarrow0$) or if there are too many errors ($FN+FP \rightarrow \infty$).  $F$ is calculated for every frame and the mark for the video is the mean over all the frames so far, $\mu(F)$.  If the predictions for all the frames were perfect, $\mu(F)$ would be equal to 1; the closest to zero, the more errors there were.  The simulation stops sometimes, if a model has become unstable; from that frame on, the value of $F$ is zero.  The model with the highest final value of $\mu(F)$ is considered the best.

@@

Additionally and independently from the mass-spring system,  it is possible to make predictions about the expected reaction force of the material, by using information from the stress-strain diagram obtained from the training data set [\fref{fig:sstrain}].  The curve in this diagram is approximated by segmented regression curves.  Several candidates are proposed using the least squares technique and the best are chosen as follows.  Since there is only one independent variable the equations for adjusting a line are straightforward:
\begin{align}
 y & = mx + y_0 + Error \\
 m & = \dfrac{n\sum x_i y_i-\sum x_i \sum y_i}{n \sum x_i^2 -(\sum x_i)^2} \\
 y_0 & = \dfrac{\sum y_i}{n} - y_0 \dfrac{\sum x_i}{n} \displaybreak[0] \\
 Error & = \sum(y_i - (mx_i+y0))^2
\end{align}
Where $y$ is the dependent variable, $x$ the independent one, $m$ is the sloop of the line, and $y_0$ the $y$ intercept.  To adjust logarithmic curves a change of variable is used.  Before computing the line, the natural logarithm of the independent variable is calculated, that is: $x'_i = \ln(x_i)$.  In this case, the final equation will be of the form: $y = m \ln x + y_0$.  It would be possible to add other types of curves, but these were enough for the materials covered.

>>>>

The contour of the deformable object is used to obtain the ground truth for the training phase, which is the area enclosed by the contour. \comment{Since we only require a tracking algorithm which is good enough to evaluate the predictions of the mass-spring system, a very simple one was used.}  Geometrically, the contour is represented by a polygon with hundreds of sides, which can also be called a \textit{linear spline} \comment{\cite{Menet1991}. The vertices of the polygon work as control points which can be moved to adjust the polygon to the shape of the object as it gets deformed.  Even though this implementation is inspired in the theory of active contours, it works with a much simpler and faster, but less accurate algorithm as it is explained below, instead of minimizing a global energy.}  The tracking of the deformable target in the ground truth has two stages:
\begin{enumerate}
 \item The target object is segmented from the environment.  The Canny algorithm is used to extract the edges of the image.  \comment{\sout{Next, a simple colour segmentation is used to remove edges that are not likely to belong to the object of interest.}  Since the object of interest has an outstanding colour, the hue of the pixels on every edge are used to keep only edges that could belong to it.  If the average hue of the pixels on an edge fall outside a previously calibrated range, the edge is eliminated.}
 \item The internal representation of its contour, the linear snake, is adapted to its new shape.  The algorithm assumes that no occlusions take place and that the contour does not bend over itself.
\end{enumerate}
For the first frame, the snake is initialised as a rectangle around the target. This rectangle fragments itself into smaller pieces. Each new vertex adheres itself to the closest point on an edge, as reached by growing rings of pixels around its current position, up to a certain threshold distance [\fref{fig:scene}(a)].  If there is no edge within that area, the control point remains were it is. This criterion was selected to make the representation more robust against edges that are easier to detect on some frames than in others.  Once the edge reappears, the control point adapts itself again.  Also, the distance between two consecutive control points is kept within the interval $[min\_distance, max\_distance]$ and two adjacent edges do not form angles smaller than $min\_angle$.  See \alref{alg:regularise}.  This allows the snake to adapt very quickly to the complexity of the contour it represents.

@@

This measurement \comment{is know as F-score and} is used in the computer vision community to evaluate segmentation algorithms.  $F=1$ means that there were no false predictions ($FN+FP = 0$).  $F \rightarrow 0$\footnote{This is read: the value of $F$ goes to zero or $F$ tends to zero.}, if there are no good predictions ($TP\rightarrow0$) or if there are too many errors ($FN+FP \rightarrow \infty$).  $F$ is calculated for every frame and the mark for the video \comment{ at frame $i$} is the mean over all the frames so far, $\mu(F)$:
\comment{\begin{align}
 \mu(F) = \sum_{t=0}^{t=i} F(frame=t)
\end{align}}
If the predictions for all the frames were perfect, $\mu(F)$ would be equal to 1; the closest to zero, the more errors there were.  The simulation stops sometimes, if a model has become unstable; from that frame on, the value of $F$ is zero.  The model with the highest final value of $\mu(F)$ is considered the best.

@@

Additionally and independently from the mass-spring system,  it is possible to make predictions about the expected reaction force of the material, by using information from the stress-strain diagram obtained from the \comment{same} training data set \comment{than the mass-spring model.  This diagram is obtained from the video recorded by the visual system of the robot and the forces recoded by a sensor located at the wrist.  When the readings of the force sensor start to increase and the figer collides with the block of material, the point of zero deformation is identified.  From here on, the displacement of the finger is considered equivalent to the amount of deformation in the material.  Even though this method is not as rigorous as an engineering test, the estimation is good enough for robot manipulation of common materials.}

The \comment{stress strain curves} in [\fref{fig:sstrain}] are approximated by segmented regression curves.  Several candidates are proposed using the least squares technique and the best are chosen as follows.  Since there is only one independent variable the equations for adjusting a line are straightforward:
\begin{align}
 y & = mx + y_0 + Error \\
 m & = \dfrac{n\sum x_i y_i-\sum x_i \sum y_i}{n \sum x_i^2 -(\sum x_i)^2} \\
 y_0 & = \dfrac{\sum y_i}{n} - y_0 \dfrac{\sum x_i}{n} \displaybreak[0] \\
 Error & = \sum(y_i - (mx_i+y0))^2
\end{align}
Where $y$ is the dependent variable, $x$ the independent one, $m$ is the sloop of the line, and $y_0$ the $y$ intercept.  To adjust logarithmic curves a change of variable is used.  Before computing the line, the natural logarithm of the independent variable is calculated, that is: $x'_i = \ln(x_i)$.  In this case, the final equation will be of the form: $y = m \ln x + y_0$.  It would be possible to add other types of curves, but these were enough for the materials covered.
